Proof of uniform convergence for a cell-centered AP discretization of the hyperbolic heat equation on general meshes

### 2015-03-20 Corrected_version.tex
Chgt constante + Lemma 2.8

Provided that the initial data is smooth, the initialization gives us 
\begin{equation}\label{init_1D}
 \| u_j^{\eps}(0)-\eps v_j(0)\|_{L^2(\mathbb{R})} \leq C h\eps .
\end{equation}

@@

\begin{equation}\label{l2}
 \|\bar{v}_{h}\|_{L^2([0,T]\times\mathbb{R})}
=
\sqrt{
\int_0^T \sum_j \Delta x_{j+\frac12} (v_{j+\frac12} )^2
}
\leq
C\|p_0\|_{L^2(\mathbb R)}, \qquad C>0.
\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

We  first 
 recall a fundamental   error estimate \cite{FV}
for the diffusion limit scheme \eqref{schemadifff}, assuming smooth initial data \eqref{eq:intip0}. 
\begin{proposition}\label{ds1d}
There exists a constant $C(T)>0$ such that
\begin{equation*}
\|\W_h^\eps-\W^\eps\|_{L^{\infty}([0,T];L^2(\mathbb{R}))} \leq
C(T)h \| \partial_{xxx}p_0 \|_{L^2(\mathbb{R})  }.
\end{equation*}
\end{proposition}

@@

 \|\partial_{xxx}p_0   \| _{L^2(\mathbb{R})} \bigg) \sqrt{\frac12 + \frac{4}{\mathcal{C}_M}\bigg( 3 + \frac{T}{\sigma}e^{T/\sigma} \bigg) }

@@

where the constant $\mathcal{C}_M$ is defined in assumption \ref{geometrie1d}. Finally, 

@@

\|\V_h^{\eps}-\W^\eps_h\|_{L^2([0,T]\times\mathbb R)} \leq
C(T)  \eps.

@@

Thanks to the estimate \eqref{l2} and as the scheme is linear, $\frac{d}{dt} u_j$

@@

\|R\|_{L^2([0,T]\times \mathbb{R})}=\bigg\|\frac{d}{dt} u_j\bigg\|_{L^2([0,T]\times \mathbb{R})}\leq
C\eps\bigg\|\frac{d}{dt} p_h(0)\bigg\|_{L^2(\mathbb{R})}.

@@

implies that $\|\frac{d}{dt} p_h(0)\|_{L^2(\mathbb{R})}\leq C 
\| p(t=0)\|_{H^2(\mathbb{R})}$. So one has the bound

@@

\|R\|_{L^2([0,T]\times \mathbb{R})}\leq C \eps
It is seen in proposition \ref{pro:H1}
that 

@@

\| S\|_{L^2(\mathbb{R})} \leq \eps  

@@

^2 \leq \frac12 \|\V_h^{\eps}-\W^\eps_h\|_{L^2(\mathbb R)}^2 +
C\bigg( \| R\|_{L^2(\mathbb R)}^2 +\| S \|_{L^2(\mathbb R)}^2\bigg).

@@

+ C (
 \| R\|_{L^2([0,T]\times\mathbb R)}^2 + \|
S\|_{L^2([0,T]\times\mathbb R)}^2
) \right)

@@

\|\V_h^{\eps}-\W^\eps_h\|_{L^2([0,T]\times\mathbb{R})} \leq C(T) \eps$. 

@@

\begin{lemma}\label{pro:H1}
The bound (\ref{eq:xx}) holds.
\end{lemma}

@@

Using the definition of $u_j$
and $u_{j+\frac12 }$, one has 
$$
S_{j+\frac12 }=\frac{\eps}{2}\left( \frac{ \Delta x_j}{\Delta x_{j+\frac12}}
\frac{d}{dt} p_j - \frac{\Delta x_{j+1}}{\Delta x_{j+\frac12}} \frac{d}{dt} p_{j+1} \right)
$$
 at all time.
Moreover, $z_h=\frac{d}{dt} p_h$ is solution of
$P_h^0$:
$$
\Delta x_j\frac{d}{dt} z_j -\frac{1}{\sigma } \bigg(
\frac{z_{j+1}-z_j}{\Delta x_{j+\frac12 }}-\frac{z_j -z_{j-1}}{\Delta
x_{j-\frac12 }}\bigg)=0,
$$
with initial condition
\begin{equation}\label{eq:zz}
z_j(0)=\frac{d}{dt} p_0(x_j)=
\frac1 {\Delta x_j \sigma } \bigg(
\frac{p_0(x_{j+1})-p_0(x_j)}{\Delta x_{j+\frac12 }}-\frac{p_0(x_j)
 -p_0(x_{j-1}  )}{\Delta
x_{j-\frac12 }}\bigg).
\end{equation}
One gets from a Taylor expansion with integral residue
that
$$
\left|
\frac{p_0( x_{j+1}) -p_0(x_j)}{\Delta x_{j+\frac12 }}- 
\partial_x p _0(x_j)
\right|\leq 
 \int_{x_j}^{x_{j+1}  } \left|\partial_{xx}p_0(y)   \right| dy .$$ 
Similarly one has the bound
$
\left|
 \frac{p_0( x_{j}) -p_0(x_{j-1})}{\Delta x_{j+\frac12 }}-
\partial_x p _0(x_j)\right|
\leq
 \int_{x_{j-1}}^{x_{j}  } \left|\partial_{xx}p_0(y)   \right| dy $.
Therefore $|z_j(0)|\leq
\frac1 {\Delta x_j \sigma }  \int_{x_{j-1}}^{x_{j+1}  } \left|\partial_{xx}p_0(y)  
 \right| dy $ from which the bound
$
\sqrt{ \sum_j \Delta x_j z_j^2(0) }
\leq \sigma^{-1}\|p_0\|_{H^2(\mathbb{R})}$
is deduced.
Since the scheme $P_h^0$ is stable in $L^2$, this bound is true at any time.
Considering (\ref{eq:zz}) the discrete second derivative
attached to $P_h^0$ is bounded at any time, which ends the proof
of the claim.
\end{proof}

>>>>

Provided that the initial data is smooth, the initialization gives us 
\begin{equation}\label{init_1D}
 \| u_j^{\eps}(0)-\eps v_j(0)\|_{L^2(\mathbb{R})} \leq C_0 h\eps .
\end{equation}

@@

\begin{equation}\label{l2}
 \|\bar{v}_{h}\|_{L^2([0,T]\times\mathbb{R})}
=
\sqrt{
\int_0^T \sum_j \Delta x_{j+\frac12} (v_{j+\frac12} )^2
}
\leq
\sqrt{\frac{\sigma}{2}}\|p_0\|_{L^2(\mathbb R)}, \qquad C>0.
\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

We  first 
 recall a fundamental   error estimate \cite{FV}
for the diffusion limit scheme \eqref{schemadifff}, assuming smooth initial data \eqref{eq:intip0}. 
\begin{proposition}\label{ds1d}
Under assumption \ref{geometrie1d}, the following estimate holds
\begin{equation*}
\|\W_h^\eps-\W^\eps\|_{L^{\infty}([0,T];L^2(\mathbb{R}))} \leq
h 
\left(  \|\partial_{xx}p_0   \| _{L^2(\mathbb{R})}+
 \|\partial_{x}p_0   \| _{L^2(\mathbb{R})}
\right)  \bigg(  
e^{T/2\sigma} + 
\eps \sqrt{\frac12 + \frac{4}{\mathcal{C}_M}\bigg( 3 + \frac{T}{\sigma}e^{T/\sigma} \bigg) } \bigg).
\end{equation*}
\end{proposition}

@@

 \|\partial_{xxx}p_0   \| _{L^2(\mathbb{R})} \bigg) \sqrt{\frac12 + \frac{4}{C_{\mathcal{M}}}\bigg( 3 + \frac{T}{\sigma}e^{T/\sigma} \bigg) }

@@

where the constant $C_{\mathcal{M}}$ is defined in assumption \ref{geometrie1d}. Finally, 

@@

\|\V_h^{\eps}-\W^\eps_h\|_{L^2([0,T]\times\mathbb R)} \leq  \eps e^{t/2} \bigg(  C_0h + \sqrt{ \frac{1}{\sigma C_{\mathcal{M}} } ( \frac12 + \frac{1}{\sigma^2} )} \|p_0\|_{H^2(\mathbb{R})} \bigg) ,
where $C_0$ is defined in \eqref{init_1D}.

@@

Thanks to the estimate \eqref{l2} and since the scheme is linear, $\frac{d}{dt} u_h = (\frac{d}{dt} u_j)_{j\in\mathbb{Z}}$

@@

\bigg\|\frac{d}{dt} u_h\bigg\|_{L^2([0,T]\times \mathbb{R})} = \sqrt{\int_0^T \sum_{j\in\mathbb{Z}} \Delta x_j u_j^2 }\leq
\eps \sqrt{ \frac{\sigma}{2C_{\mathcal{M}} }} \bigg\|\frac{d}{dt} p_h(0)\bigg\|_{L^2(\mathbb{R})}.

@@

implies that $\|\frac{d}{dt} p_h(0)\|_{L^2(\mathbb{R})}\leq \frac{1}{\sigma}
\| p_0\|_{H^2(\mathbb{R})}$  (proposition \ref{pro:H1}). So one has the bound

@@

\|R\|_{L^2([0,T]\times \mathbb{R})}\leq   \frac{\eps }{ \sqrt{ 2\sigma C_{\mathcal{M}} } } 
\| p_0\|_{H^2(\mathbb{R})}.
Using the definitions of $u_j$ \eqref{schemadiff3}, the $L^2$ norm of $S_h = (S_{j+\frac12})_{j\in\mathbb{Z}}$ is related to the $L^2$ norm of $\frac{d}{dt}u_j$. Using once again the proposition \ref{pro:H1}, one finds out that 

@@

\| S_h \|_{L^2(\mathbb{R})} \leq \eps  \frac{\sqrt{2}}{\sigma \sqrt{C_{\mathcal{M}}}}

@@

 \frac{d}{dt}\|\V_h^{\eps}-\W^\eps_h\|_{L^2(\mathbb R)}
^2 \leq  \|\V_h^{\eps}-\W^\eps_h\|_{L^2(\mathbb R)}^2 +
 \| R\|_{L^2(\mathbb R)}^2 + \frac1{2\sigma}\| S \|_{L^2(\mathbb R)}^2 . 

@@

 \| R\|_{L^2([0,T]\times\mathbb R)}^2 +  \frac1{2\sigma} \|
S\|_{L^2([0,T]\times\mathbb R)}^2 \right)

@@

\|\V_h^{\eps}-\W^\eps_h\|_{L^2([0,T]\times\mathbb{R})} \leq \eps e^{t/2} \big(  C_0h + \sqrt{ \frac{1}{\sigma C_{\mathcal{M}} } ( \frac12 + \frac{1}{\sigma^2} )} \|p_0\|_{H^2(\mathbb{R})} \big)$. 

@@

\begin{lemma}\label{pro:H1}
The bound $
\sqrt{ \sum_j \Delta x_j (\frac{d}{dt}u_j)^2(t) }
\leq \sigma^{-1}\|p_0\|_{H^2(\mathbb{R})}$
 holds at any time.\end{lemma}

@@

By linearity of the diffusion scheme, $z_h=\frac{d}{dt} p_h$ is solution of
$P_h^0$:
$$
\Delta x_j\frac{d}{dt} z_j -\frac{1}{\sigma } \bigg(
\frac{z_{j+1}-z_j}{\Delta x_{j+\frac12 }}-\frac{z_j -z_{j-1}}{\Delta
x_{j-\frac12 }}\bigg)=0,
$$
with initial condition
\begin{equation}\label{eq:zz}
z_j(0)=\frac{d}{dt} p_0(x_j)=
\frac1 {\Delta x_j \sigma } \bigg(
\frac{p_0(x_{j+1})-p_0(x_j)}{\Delta x_{j+\frac12 }}-\frac{p_0(x_j)
 -p_0(x_{j-1}  )}{\Delta
x_{j-\frac12 }}\bigg).
\end{equation}
One gets from a Taylor expansion with integral residue
that
$$
\left|
\frac{p_0( x_{j+1}) -p_0(x_j)}{\Delta x_{j+\frac12 }}- 
\partial_x p _0(x_j)
\right|\leq 
 \int_{x_j}^{x_{j+1}  } \left|\partial_{xx}p_0(y)   \right| dy .$$ 
Similarly one has the bound
$
\left|
 \frac{p_0( x_{j}) -p_0(x_{j-1})}{\Delta x_{j+\frac12 }}-
\partial_x p _0(x_j)\right|
\leq
 \int_{x_{j-1}}^{x_{j}  } \left|\partial_{xx}p_0(y)   \right| dy $.
Therefore $|z_j(0)|\leq
\frac1 {\Delta x_j \sigma }  \int_{x_{j-1}}^{x_{j+1}  } \left|\partial_{xx}p_0(y)  
 \right| dy $ from which the bound
$
\sqrt{ \sum_j \Delta x_j z_j^2(0) }
\leq \sigma^{-1}\|p_0\|_{H^2(\mathbb{R})}$
is deduced.
Since the scheme $P_h^0$ is stable in $L^2$, this bound is true at any time.
Considering (\ref{eq:zz}) the discrete second derivative
attached to $P_h^0$ is bounded at any time, which ends the proof
of the claim.
\end{proof}

### 2015-03-20 U_v16.tex
Chgt constante + Lemma 2.8

then  in terms of $p^{\eps} $ an $u^{\eps} $ we have evidently

@@

\begin{equation}\label{l2}
 \|\bar{v}_{h}\|_{L^2([0,T]\times\mathbb{R})}
=
\sqrt{
\int_0^T \sum_j \Delta x_{j+\frac12} (v_{j+\frac12} )^2
}
\leq
C\|p_0\|_{L^2(\mathbb R)}, \qquad C>0.
\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{proposition}\label{pro:H1}
The bound (\ref{eq:xx}) holds.
\end{proposition}

@@

Using the definition of $u_j$
and $u_{j+\frac12 }$, one has 
$$
S_{j+\frac12 }=\frac{\eps}{2}\left( \frac{ \Delta x_j}{\Delta x_{j+\frac12}}
\frac{d}{dt} p_j - \frac{\Delta x_{j+1}}{\Delta x_{j+\frac12}} \frac{d}{dt} p_{j+1} \right)
$$
 at all time.
Moreover, $z_h=\frac{d}{dt} p_h$ is solution of
$P_h^0$:
$$
\Delta x_j\frac{d}{dt} z_j -\frac{1}{\sigma } \bigg(
\frac{z_{j+1}-z_j}{\Delta x_{j+\frac12 }}-\frac{z_j -z_{j-1}}{\Delta
x_{j-\frac12 }}\bigg)=0,
$$
with initial condition
\begin{equation}\label{eq:zz}
z_j(0)=\frac{d}{dt} p_0(x_j)=
\frac1 {\Delta x_j \sigma } \bigg(
\frac{p_0(x_{j+1})-p_0(x_j)}{\Delta x_{j+\frac12 }}-\frac{p_0(x_j)
 -p_0(x_{j-1}  )}{\Delta
x_{j-\frac12 }}\bigg).
\end{equation}
One gets from a Taylor expansion with integral residue
that
$$
\left|
\frac{p_0( x_{j+1}) -p_0(x_j)}{\Delta x_{j+\frac12 }}- 
\partial_x p _0(x_j)
\right|\leq 
 \int_{x_j}^{x_{j+1}  } \left|\partial_{xx}p_0(y)   \right| dy .$$ 
Similarly one has the bound
$
\left|
 \frac{p_0( x_{j}) -p_0(x_{j-1})}{\Delta x_{j+\frac12 }}-
\partial_x p _0(x_j)\right|
\leq
 \int_{x_{j-1}}^{x_{j}  } \left|\partial_{xx}p_0(y)   \right| dy $.
Therefore $|z_j(0)|\leq
\frac1 {\Delta x_j \sigma }  \int_{x_{j-1}}^{x_{j+1}  } \left|\partial_{xx}p_0(y)  
 \right| dy $ from which the bound
$
\sqrt{ \sum_j \Delta x_j z_j^2(0) }
\leq \sigma^{-1}\|p_0\|_{H^2(\mathbb{R})}$
is deduced.
Since the scheme $P_h^0$ being stable in $L^2$, this bound is true at any time.
Considering (\ref{eq:zz}) the discrete second derivative
attached to $P_h^0$ is bounded at any time, which ends the proof
of the claim.
\end{proof}

>>>>

then  in terms of $p^{\eps} $ and $u^{\eps} $ we have evidently

@@

\begin{equation}\label{l2}
 \|\bar{v}_{h}\|_{L^2([0,T]\times\mathbb{R})}
=
\sqrt{
\int_0^T \sum_j \Delta x_{j+\frac12} (v_{j+\frac12} )^2
}
\leq
\sqrt{\frac{\sigma}{2}} \|p_0\|_{L^2(\mathbb R)}.
\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{proposition}\label{pro:H1}
The bound $
\sqrt{ \sum_j \Delta x_j (\frac{d}{dt}u_j)^2(0) }
\leq \sigma^{-1}\|p_0\|_{H^2(\mathbb{R})}$
 holds at any time.
\end{proposition}

@@

By linearity of the diffusion scheme, $z_h=\frac{d}{dt} p_h$ is solution of
$P_h^0$:
$$
\Delta x_j\frac{d}{dt} z_j -\frac{1}{\sigma } \bigg(
\frac{z_{j+1}-z_j}{\Delta x_{j+\frac12 }}-\frac{z_j -z_{j-1}}{\Delta
x_{j-\frac12 }}\bigg)=0,
$$
with initial condition
\begin{equation}\label{eq:zz}
z_j(0)=\frac{d}{dt} p_0(x_j)=
\frac1 {\Delta x_j \sigma } \bigg(
\frac{p_0(x_{j+1})-p_0(x_j)}{\Delta x_{j+\frac12 }}-\frac{p_0(x_j)
 -p_0(x_{j-1}  )}{\Delta
x_{j-\frac12 }}\bigg).
\end{equation}
One gets from a Taylor expansion with integral residue
that
$$
\left|
\frac{p_0( x_{j+1}) -p_0(x_j)}{\Delta x_{j+\frac12 }}- 
\partial_x p _0(x_j)
\right|\leq 
 \int_{x_j}^{x_{j+1}  } \left|\partial_{xx}p_0(y)   \right| dy .$$ 
Similarly one has the bound
$
\left|
 \frac{p_0( x_{j}) -p_0(x_{j-1})}{\Delta x_{j+\frac12 }}-
\partial_x p _0(x_j)\right|
\leq
 \int_{x_{j-1}}^{x_{j}  } \left|\partial_{xx}p_0(y)   \right| dy $.
Therefore $|z_j(0)|\leq
\frac1 {\Delta x_j \sigma }  \int_{x_{j-1}}^{x_{j+1}  } \left|\partial_{xx}p_0(y)  
 \right| dy $ from which the bound
$
\sqrt{ \sum_j \Delta x_j z_j^2(0) }
\leq \sigma^{-1}\|p_0\|_{H^2(\mathbb{R})}$
is deduced.
Since the scheme $P_h^0$ being stable in $L^2$, this bound is true at any time.
Considering (\ref{eq:zz}) the discrete second derivative
attached to $P_h^0$ is bounded at any time, which ends the proof
of the claim.
\end{proof}
